Supporting Information for: A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal

### 2020-09-01 Manuscript/SupportingInfo.Rmd
z12 updates

Here we examine the statistical properties of tests for comparing the relative strength of phylogenetic signal, using $\hat{Z}_{12}$. Simulations were performed using pure-birth trees of different sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$). For each combination of $n$ and $\lambda$, 101 traits were simulated on the phylogeny under a Brownian motion model of evolution. Next, the phylogenetic signal of trait 1 was compared to the signal in the remaining 100 traits in pairwise fashion, using $\hat{Z}_{12}$. The proportion of significant results provided an estimate of the type I error or model misspecification. Specifically, type I error was evaluated when traits were simulated using no input phylogenetic signal (i.e., $\lambda = 0$). Likewise, model misspecification was evaluated for data simulated with some known, but equal, level of phylogenetic signal (i.e., $\lambda > 0$). \hfill\break

**Results:** Tests revealed that across simulation conditions, the average type I error of $\hat{Z}_{12}$ was approximately 0.05 (Fig. S21). Levels of misspecification were slightly higher, and ranged on average between 7% (when comparing traits generated on the same phylogeny: Fig. S21) to ~12% (for comparisons of traits from different trees). Both results demonstrate that comparisons of the strength of phylogenetic signal using $\hat{Z}_{12}$ do not display high levels of false positives, and are thus appropriate for comparing the degree of phylogenetic signal across traits.

@@

\textbf{Figure S21}. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$).

>>>>

Here we examine the statistical properties of tests for comparing the relative strength of phylogenetic signal, using $\hat{Z}_{12}$. Simulations were performed using pure-birth trees of different sizes ($n=2^5, 2^6, \cdots, 2^{10}$), and with differing levels of phylogenetic signal ($\lambda=0.0, 0.5, \cdots, 1.0$). For each combination of $n$ and $\lambda$, 50 traits were simulated on the phylogeny under a Brownian motion model of evolution. Next, the phylogenetic signal between traits was compared in pairwise fashion for all $50*49/2 = 1225$ combinations, using $\hat{Z}_{12}$. The proportion of significant results provided an estimate of the type I error or model misspecification. Specifically, type I error was evaluated when traits were simulated using no input phylogenetic signal (i.e., $\lambda = 0$). Likewise, model misspecification was evaluated for data simulated with some known, but equal, level of phylogenetic signal (i.e., $\lambda > 0$). \hfill\break

**Results:** Tests revealed that across simulation conditions, the average type I error of $\hat{Z}_{12}$ was approximately 0.05 (Fig. S21). Levels of misspecification were slightly higher, and ranged on average between 7% when comparing traits generated on the same phylogeny (Fig. S21 black line) to ~12% for comparisons of traits from different trees. Both results demonstrate that comparisons of the strength of phylogenetic signal using $\hat{Z}_{12}$ do not display high levels of false positives, and are thus appropriate for comparing the degree of phylogenetic signal across traits.

@@

\textbf{Figure S21}. Type I error rates and model misspecification for $\hat{Z}_{12}$. Type I error is found on the far-left of the plot ($\lambda = 0$). The solid black line represents the average type I error and false discovery rate across all levels of input signal.
